Lyman-$\alpha$ photons through rotating outflows

### 2018-03-31 paper/paper.tex
edit

\subsubsection{Kinematic Maps}
The first is related to the capability of spatially resolving the
extent of a LAE.
For instance \cite{Prescott14} presented observational results of a
Doppler shift when taking spectra at two oposite sides of a LAE.

In Figure \ref{fig:doppler} we present a toy model (\vrot$=50\ \kms$,
\vout$=25\ \kms$ and \tauh$=10^5$) for the spectrum of
a LAE taken from to different sides of the galaxy. 
As the LAE is rotating, one side is being redshifted while the other
is blueshifted. 
We see that the full spectrum is a weighted line, in solid black,
that is found between these two.
We notice that the distance between the maxima of the blue and red
spectra is not twice the rotational velocity as it could be naively
expected.
Although is a good approximation to model the rotating spectra by a
sum of Doppler shifts, the peak of the spectra is also weighted by the
amount of mass with a given velocity. 
In this toy model the distance between the peaks of the
receeding/approaching spectra is close to $\sim 20\ \kms$, which is
less than half of the naively expected value of $2\vrot=100\ \kms$

New spectrographs like MUSE should help obtain kinematic information
from LAEs to build velocity maps in \lya. 
This could be a natural extension of the work reported by
\cite{Herenz2016} on the velocity maps of several LARS (Lyman Alpha
Reference Sample) galaxies.

@@

On the other hand, regarding the parameters \vrot, \vout, and $\theta$ for
which this model would work, we have the following. The effect of the outflows
velocity is much more sensitive on the spectrum than the rotational one; meaning
that for a high \vout, the \vrot value would have to be equal or greater in order
to be able to glimpse the shift in frequency that rotation induces. Regarding $\theta$,
the introduction of rotation in the model implies that it will only affect galaxies
with its viewing angle $\theta \neq 90^\circ$. This can be observed in the cartesian
components of the velocity of the atoms (Eq. \ref{eq:vx}, \ref{eq:vy}, \ref{eq:vz})
where the only velocity factor affecting $v_z$ is the outflows one.

>>>>

\subsubsection{\lya Kinematic Maps}
The first is related to the capability of spatially resolving the
extent of a LAE.
For instance \cite{Prescott14} presented observational results of a
Doppler shift when taking spectra at two oposite sides of a LAE.

In Figure \ref{fig:doppler} we present a toy model (\vrot$=50\ \kms$,
\vout$=25\ \kms$ and \tauh$=10^5$) for the spectrum of
a LAE taken from to different sides of the galaxy. 
As the LAE is rotating, one side is being redshifted while the other
is blueshifted. 
We see that the full spectrum is a weighted line, in solid black,
that is found between these two.
We notice that the distance between the maxima of the blue and red
spectra is not twice the rotational velocity as it could be naively
expected.

Although it is a good approximation to think the rotating spectra by a
sum of Doppler shifts, the peak of the spectra is also weighted by the
amount of mass with a given velocity. 
In this toy model the distance between the peaks of the
receeding/approaching spectra is close to $\sim 20\ \kms$, which is
less than half of the naively expected value of $2\vrot=100\ \kms$,
due to the fact that only a small fraction of the photons are emitted
at the extreme of the galaxy having the maximum rotational velocity of
$50\ \kms$.

New spectrographs like MUSE should help obtain kinematic information
from LAEs to build velocity maps in \lya. 
This could be a natural extension of the work reported by
\cite{Herenz2016} on the velocity maps of several LARS (Lyman Alpha
Reference Sample) galaxies.  
The interpretation of such data should take into account the insights
and trends we have presented in this paper.

\subsubsection{\lya continuum escape or rotation?}

@@
